A Cross-Domain Benchmark for Active Learning

### 2024-06-04 doc/al_benchmark_paper.tex
paper

% TODOs nach Priorität
% !! "Overfitting on test" für Oracle: Sanity-Check durchführen und eventuell komplett streichen
% !! Introduction neu schreiben: Wir brauchen eine klare Benennung der Ziele und bessere Abgrenzung zu anderen papern 
% !! Sec 4 Methodology ist sehr chaotisch. Wir müssen neu sortieren und unwichtige Dinge streichen
% Contributions neu sortieren und ausführlicher beschreiben
% Methode zum CD-Diagramm genau erklären (paired-t-test etc.) und die Relevanz fürs paper zeigen
% Advantage-Matrix von Ji et. al. neu aufarbeiten. Wo sind die Unterschiede? Ist das CD-Diagramm besser/schlechter?
% (Un-)Encoded durch Normal uns Semi-Supervised ersetzen, um die Verbindung klar zu machen
% Gibt es eine Conclusion für die Encoded-Datasets? Im Moment werden die nicht besprochen
% Eventuell Batchsize 1 von den anderen Ergebnissen separieren. Der Vergleich macht mehr Sinn, als einfach die Batchsizes zu mitteln
% Tabelle 1 für Algorithmen statt Domänen. Das zeigt, dass immer entweder BADGE oder Margin (unsere Top 2) fehlt
% Oracle Performance mit einer Zahl benennen: z.B. Da wird ~20% lift zu allen Baselines haben, brauchen wir kein aufwendigeres Oracle
% Unsere Rankings mit Literatur vergleichen. Wie anders sind wir? -> Evtl. besten Alg. pro Benchmark sammeln und vergleichen
% Causalitäten checken: Viele Dinge werden einfach genannt, ohne das der Grund richtig eingeführt wird
% "reproducible" als Aufhänger für 50 repeats ist nicht intuitiv
% Fig. 1: Schlechte Performance am Anfang verwirrt nur. Entweder BADGE benutzen, anders plotten, oder im Text in Kontext setzen
% Eventuell Honeypot und Diverging Sine direkt in den Contributions nennen
% Oracle Forecast im Appendix erklären 
% Tabelle mit AUCs in den Appendix

%%%%%%%%%%%%%%%%%%
% DONE
%%%%%%%%%%%%%%%%%
% !! Ergebnis-Tabelle mit nur 3 runs neu berechnen und Unterschiede zeigen
% !! Least-Confident Baseline
% Conclusion: Es gibt Unterschiede bei der Top 1 zwischen den Domänen, aber die Top 3 ist konstant
% !! Sec 3 Related Work sollte Ji et al. im Detail einführen, weil wir uns an deren Guidelines halten. Alles andere sollte dort gekürzt werden
% Infos und Ergebnisse für die syntetischen Datensätze in ein Kapitel bündeln und direkt auf die Contribution beziehen

@@

% \section{Introduction-Old Backup}
% Deep neural networks (NN) have produced state-of-the-art results on many important supervised learning tasks.
% Since Deep NNs usually require large amounts of labeled training data, Active Learning (AL) can be used to instead select the most informative samples out of a large pool of unlabeled data, so that only these samples need to be labeled.
% It has been shown that a small labeled set of this nature can be used to train well-performing models \cite{beck2021effective, hu2021towards, li2022empirical, zhou2021towards}. \\
% On top of providing a principled way to label unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
% %Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
% In the last decade, many different algorithms for AL have been proposed.
% Even though, almost every method has reported lifts over all its predecessors,
% \footnote{Out of all considered algorithms for this paper, only BALD \cite{gal2017deep} did not claim a new SOTA performance in their result section.}
% AL research faces four central difficulties:
% {\let\thefootnote\relax\footnotetext{Code available at: anonymous}}
% %\let\thefootnote\relax\footnotetext{Code available at: \url{https://github.com/wernerth94/comparable-active-learning}}
% (i) The experiments are often carried out on different datasets and model architectures, hindering direct comparison,
% (ii) generalize poorly across different domains,
% (iii) the reported results can be subject to very high variance across restarts and
% (iv) are not always compared against important baselines like margin sampling \cite{wang2014new}.
% While multiple benchmark suites have been proposed to solve (i), to the best of our knowledge, we are the first to report results on all 3 data domains of tabular, image and text.
% Additionally, we provide synthetic datasets to highlight principled shortcoming of existing AL algorithms.
% Regarding (ii) and (iii), \cite{zhou2021towards} has pointed out severe inconsistencies in results of AL papers in recent years. 
% They conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm (\cite{zhou2021towards} Appendix A).
% These outlined issues lead to an inconclusive landscape of AL algorithms, where the vast majority of reported lifts are neither statistically significant, nor prove to be generalizable.
% This makes it very hard to identify the best AL algorithm, or even identifying state-of-the-art algorithms.
% In this work we propose an evaluation protocol that was designed to handle the high variance in the performances of AL algorithms as well as being fully controllable regardless of the combination of dataset, model and AL algorithm. 
% We base our work largely on \cite{ji2023randomness}, following their guidelines for a reliable evaluation of AL algorithms, while extending the number of evaluated data domains from 1 to 5. \\
% We focus on pool-based AL where a pool of unlabeled samples is fixed at the start of each experiment and one or more samples are chosen sequentially.
% In addition to the default scenario of selecting a batch of samples every iteration we incorporate the single sample case into our benchmark.
% Batched algorithms (and benchmarks) do not have a principled advantage over single-sample AL except for speed of computation.
% The problem of optimizing a portfolio of unlabeled samples in each iteration is more complicated to solve and the algorithms have systematically less information per sample to work with leading to a generally worse performance, that impacts some algorithms more than others. 
% %In Fig. \ref{fig:batch_vs_single} we present a comparison of Badge \cite{ashdeep} and Coreset\cite{sener2017active}, in the batch and single-sample setting.
% %Even though both algorithms were designed for batch sampling, both perform better in the single-sample setting.
% We propose single-sample AL as an important tool to identify the best acquisition function, rather than the best combination of acquisition function and diversity regularization. \\
% %\begin{figure}
% %    \centering
% %    \includegraphics[width=\linewidth]{img/ablation_batch_single}
% %    \caption{Comparison of single-sample vs. batch selection for two acquisition functions that were specifically designed for batch selection. Even when designed for batch selection, acquisition functions tend to perform better in the single sample setting.}
% %    \label{fig:batch_vs_single}
% %\end{figure}
% Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature. 
% We offer our datasets in two versions - normal and encoded.
% An encoded dataset was pre-encoded by an self-supervised encoder model, providing a different use case for active learning.
% Our synthetic and text datasets (which use word embeddings in the normal setting) are exempt from this, bringing our total dataset count to 14 across 5 domains (Tabular, Image, Text, Synthetic, Encoded).
% \begin{table}
% 	\centering
%         \caption{Comparison of our benchmark with the existing literature. Oracle curves serve as an approximation of the best possible AL algorithm. Including the encoded versions of our datasets we reach 14 datasets. "Semi" indicates whether the paper is employing any form of self- or semi-supervised learning. A "-" for repetitions means that we could not determine how often each experiment is repeated in the respective framework.}
% 	\label{tab:benchmark_comparison}
% 	\resizebox{\columnwidth}{!}{
% \begin{tabular}{lccccllllcl}
% Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
% \multicolumn{1}{l|}{Beck et al. \cite{beck2021effective}} & batch & 4 & 7 & \checkmark & - & - & - & - & - & - \\
% \multicolumn{1}{l|}{Hu et al. \cite{hu2021towards}} & batch & 5 & 13 & \checkmark & \checkmark & - & - & - & - & 3 \\
% \multicolumn{1}{l|}{Zhou et al. \cite{zhou2021towards}} & batch & 3 & 2 & \checkmark & \checkmark & - & - & - & \checkmark & 5 \\
% \multicolumn{1}{l|}{Zhan et al. \cite{zhan2021comparative}} & sngl+batch & 35 & 18 & - & - & \checkmark & \checkmark & - & \checkmark & 10-100 \\
% \multicolumn{1}{l|}{Munjal et al. \cite{munjal2022towards}} & batch & 2 & 8 & \checkmark & - & - & - & - & - & 3 \\
% \multicolumn{1}{l|}{Li et al. \cite{li2022empirical}} & batch & 5 & 13 & \checkmark & - & - & - & \checkmark & - & - \\

% \multicolumn{1}{l|}{Rauch et al. \cite{rauch2023activeglae}} & batch & 11 & 5 & - & \checkmark & - & - & - & - & 5 \\
% \hdashline
% \multicolumn{1}{l|}{Ji et al. \cite{ji2023randomness}} & batch & 3 & 8 & \checkmark & - & - & - & - & - & - \\
% \multicolumn{1}{l|}{Lueth et al. \cite{luth2024navigating}} & batch & 4 & 5 & \checkmark & - & - & - & \checkmark & - & 3 \\
% \multicolumn{1}{l|}{\textbf{Ours}} & sngl+batch & 9(14) & 11 & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & 50 
% \end{tabular}
% }
% \end{table}

% \subsection*{Contributions}
% %Maybe absorb this into the introduction text
% \begin{enumerate}
% 	\item[C1] Study on the reproducibility of AL experiments, providing evidence that previous works might have not repeated their experiments often enough to provide reliable results
% 	\item[C2] Efficient and performant algorithm for an oracle that can be constructed greedily and does not rely on search
% 	\item[C3] Two novel synthetic datasets that highlight principled shortcomings of of the top-performing AL algorithms
%     \item[C4] Extending the evaluation of Active Learning algorithms to 5 data domains and revealing significant differences in algorithm performance between them
% \end{enumerate}

@@

To sum up, \emph{ALBench} is an experimental framework which includes an
efficient oracle approximation, multiple application domains, enough repetitions to
draw valid conclusions and two synthetic tasks to highlight shortcomings of AL methods. 
By being the first benchmark to proving these things in one code-base, 
we believe that \emph{ALBench} is a major step forward of assessing the overall state of Active Learning research, independent of
specific application domains. \emph{ALBench} is publicly available under \emph{anonymous}. \\ [1mm]
Our contributions include the following:

\begin{enumerate}
\item We show that the small of repetitions that previous works have employed is not
sufficient for meaningful conclusions. Sometimes even making it impossible to assess if a performance is above or below random.
\item We propose an efficient and performant oracle which is constructed iteratively in a greedy fashion,
overcoming major computational hurdles.
% one of the major obstacles for sufficient repetitions of AL experiments. 
% [Thorben] Das können wir so nicht sagen, glaube ich. Niemand verlangt von uns, dass wir das Oracle auch 50 mal wiederholen
\item We propose \emph{IMTSBench}, the first general benchmark providing tasks
in the domains of images, text and tabular learning. It further contains
synthetic and pre-encoded data to allow for a sophisticated evaluation of AL
methods. Our experiments show, that there is no clear state-of-the art method
for active learning across different domains.
\item We propose \emph{Honeypot} and \emph{Diverging Sin}, two synthetic datasets
designed to hinder active learning by naive decision-boundary- or
clustering-based approaches respectively. Thus, they provide an important tool to identify 
shortcomings of existing AL methods.
\end{enumerate}

%\todo{Ich bin mir nicht ganz sicher, wie man Honeypot und Diverging Sin verkauft. Macht as mit der Lower bar sinn?}
% \begin{enumerate}
% \item[C1] Study on the reproducibility of AL experiments, providing evidence
% that previous works might have not repeated their experiments often enough to
% provide reliable results
% \item[C2] Efficient and performant algorithm for an oracle that can be
% constructed greedily and does not rely on search
% \item[C3] Two novel synthetic datasets that highlight principled shortcomings of
% of the top-performing AL algorithms
% \item[C4] Extending the evaluation of Active Learning algorithms to 5 data
% domains and revealing significant differences in algorithm performance between
% them to overcome one of the major obstacles for sufficient 
% \end{enumerate}

% we are the first to report results on all 3 data domains of
% tabular, image and text. Additionally, we provide synthetic datasets to
% highlight principled shortcoming of existing AL algorithms. Regarding (ii) and
% (iii), \cite{zhou2021towards} has pointed out severe inconsistencies in results
% of AL papers in recent years. They conducted a meta analysis of reported results
% of several different AL algorithms and found that all considered algorithms only
% provided significant lifts in their own original papers, while following
% literature reported performances no better that uncertainty sampling, or in some
% cases no better than random sampling for the same algorithm
% (\cite{zhou2021towards} Appendix A). These outlined issues lead to an
% inconclusive landscape of AL algorithms, where the vast majority of reported
% lifts are neither statistically significant, nor prove to be generalizable. This
% makes it very hard to identify the best AL algorithm, or even identifying
% state-of-the-art algorithms. In this work we propose an evaluation protocol that
% was designed to handle the high variance in the performances of AL algorithms as
% well as being fully controllable regardless of the combination of dataset, model
% and AL algorithm. We base our work largely on \cite{ji2023randomness}, following
% their guidelines for a reliable evaluation of AL algorithms, while extending the
% number of evaluated data domains from 1 to 5. \\
% We focus on pool-based AL where a pool of unlabeled samples is fixed at the
% start of each experiment and one or more samples are chosen sequentially. In
% addition to the default scenario of selecting a batch of samples every iteration
% we incorporate the single sample case into our benchmark. Batched algorithms
% (and benchmarks) do not have a principled advantage over single-sample AL except
% for speed of computation. The problem of optimizing a portfolio of unlabeled
% samples in each iteration is more complicated to solve and the algorithms have
% systematically less information per sample to work with leading to a generally
% worse performance, that impacts some algorithms more than others. 
% %In Fig. \ref{fig:batch_vs_single} we present a comparison of Badge \cite{ashdeep} and Coreset\cite{sener2017active}, in the batch and single-sample setting.
% %Even though both algorithms were designed for batch sampling, both perform better in the single-sample setting.
% We propose single-sample AL as an important tool to identify the best
% acquisition function, rather than the best combination of acquisition function
% and diversity regularization. \\
%\begin{figure}
%    \centering
%    \includegraphics[width=\linewidth]{img/ablation_batch_single}
%    \caption{Comparison of single-sample vs. batch selection for two acquisition functions that were specifically designed for batch selection. Even when designed for batch selection, acquisition functions tend to perform better in the single sample setting.}
%    \label{fig:batch_vs_single}
%\end{figure}
% Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature. 
% We offer our datasets in two versions - normal and encoded.
% An encoded dataset was pre-encoded by an self-supervised encoder model, providing a different use case for active learning.
% Our synthetic and text datasets (which use word embeddings in the normal
% setting) are exempt from this, bringing our total dataset count to 14 across 5
% domains (Tabular, Image, Text, Synthetic, Encoded). 
\begin{table}
\caption{Comparison of our benchmark with the existing literature. Oracle curves
serve as an approximation of the best possible AL algorithm. Including the
encoded versions of our datasets we reach 14 datasets, without we have 9. "Semi"
indicates whether the paper is employing any form of self- or semi-supervised
learning. A "-" for repetitions means that we could not determine how often each
experiment is repeated in the respective framework. \emph{ALBench} is the only
benchmark which reports over enough runs for meaningful results and considers
all 5 domains.}
	\label{tab:benchmark_comparison}
\begin{tabular}{lccc|cllll|cl}
Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
\multicolumn{1}{l|}{Beck et al. \cite{beck2021effective}} & batch & 4 & 7 & \checkmark & - & - & - & - & - & - \\
\multicolumn{1}{l|}{Hu et al. \cite{hu2021towards}} & batch & 5 & 13 & \checkmark & \checkmark & - & - & - & - & 3 \\
\multicolumn{1}{l|}{Zhou et al. \cite{zhou2021towards}} & batch & 3 & 2 & \checkmark & \checkmark & - & - & - & \checkmark & 5 \\
\multicolumn{1}{l|}{Zhan et al. \cite{zhan2021comparative}} & sngl+batch & 35 & 18 & - & - & \checkmark & \checkmark & - & \checkmark & 10-100 \\
\multicolumn{1}{l|}{Munjal et al. \cite{munjal2022towards}} & batch & 2 & 8 & \checkmark & - & - & - & - & - & 3 \\
\multicolumn{1}{l|}{Li et al. \cite{li2022empirical}} & batch & 5 & 13 & \checkmark & - & - & - & \checkmark & - & - \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
While multiple benchmark suites have been proposed for Active Learning, none of them provide experiments for more than two domains.
The authors of \cite{beck2021effective}, \cite{munjal2022towards}, \cite{li2022empirical}, \cite{ji2023randomness} and \cite{luth2024navigating} even focus exclusively on the image domain.
Especially the tabular domain is underrepresented in preceding benchmarks, as only \cite{zhan2022comparative} provides experiments for it.
Experiments on the interplay between AL and semi-supervised learning are similarly under-researched, as only two works exist \cite{li2022empirical, luth2024navigating}, both of them only using images.
An oracle algorithm has been proposed by two works \cite{zhou2021towards, zhan2022comparative}. 
Both of these algorithms rely on search and are computationally very expensive, while our proposed method efficiently can be constructed sequentially.
%While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training, and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning.
%The authors of \cite{hu2021towards} study models that are trained on actively learned datasets in the image and text domain.
%They test for several different properties of the models including robustness, response to compression techniques and final performance.
%\cite{zhou2021towards} proposed an oracle algorithm for AL that uses Simulated Annealing search to approximate a solution for the optimal subset of labeled data.
%Additionally, they study the generalization behavior of subsets of labeled data in the text an image domain.
The two closest related works to this benchmark are \cite{ji2023randomness} and \cite{luth2024navigating}, who also place a much higher emphasis on the problem of evaluating AL algorithms under many forms of variance than their predecessors (indicated in Tab. \ref{tab:benchmark_comparison} by a dashed line).
The authors of \cite{ji2023randomness} posed a total of 12 "recommendations" for reliable evaluation of AL algorithms.
We largely adapt the proposed recommendations of \cite{ji2023randomness} and extend their work to multiple domains, batch sizes and comparisons.
For a complete list of the recommendations and our implementation of them, please refer to App. \ref{app:recommendations}.
This work also pays attention to the so-called "pitfalls" of AL evaluation proposed in \cite{luth2024navigating}.
For a complete list of the pitfalls and our implementation of them, please refer to App. \ref{app:pitfalls}.
To the best of our knowledge, we are the first to extend reliable SOTA (based on \cite{ji2023randomness, luth2024navigating}) experimentation to a total of 5 data domains and a high number of restarts per experiment.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Few Repetitions are not Sufficient for Meaningful Results}\label{sec:restarts}
To evaluate how many restarts are necessary to obtain conclusive results in an AL experiment, we computed 100 runs of our top-performing algorithm on one dataset.
Our best algorithm is margin sampling and we chose the Splice dataset for its average size and complexity. \\
This allows us firstly, to obtain a very strong estimation of the "true" average performance of margin sampling on this dataset and secondly, to draw subsets from this pool of 100 runs.
Setting the size of our draws to $\alpha$ and sampling uniformly, we can approximate a cross-validation process with $\alpha$ restarts.
Each of these draws can be interpreted as a \textbf{reported result in AL literature} where the authors employed $\alpha$ restarts.
%We provide evidence in Fig. \ref{fig:restarts} that previous works might not have evaluated their experiments with a sufficient number of restarts.
Figure \ref{fig:restarts} shows the "true" mean performance of margin sampling (green) in relation to random sampling (black) and the oracle performance (red).
We display 5 random draws of size $\alpha$ in blue.
We can observe that even for a relatively high number of restarts the variance between the samples is extremely high, resulting in some performance curves being worse that random and some being significantly better.
When setting $\alpha = 50$ we observe all samples to converge close to the true mean performance. 
\begin{figure}
        \hspace{-20mm}
	\includegraphics[width=1.24\linewidth]{img/ablation_restarts}
	\caption{Random draws from a pool of 100 runs for margin sampling on the Splice dataset with different numbers of repetitions ($\alpha=\{5,10,50\}$). Green curves are the mean performance of all 100 runs, while the samples are blue. Even with 5 or 10 repetitions, we can observe that single draws for margin sampling display below-random performance (black), while the true mean should be above random.}
	\label{fig:restarts}
\end{figure}
In addition to this motivating example, we carried out our main evaluation (Tab. \ref{tab:results}) multiple times by uniformly sampling 3 random from our available runs and comparing the results.
We found significant differences in the performance of acquisition functions on individual datasets, as well as permutations in the final ranking.
This partly explains the ongoing difficulties in reproducing results for AL experiments and benchmarks.
The details can be found in App. \ref{app:rank_difference}.
For this benchmark we employ 50 restarts of every experiment.

>>>>

% TODOs nach Priorität
% Terminologie aufräumen:
% - Query size vs batch size
% - Acq. function vs. AL Alg. vs. AL Method
% - (Un-)encoded vs. semi-supervised/normal
% Gibt es eine Conclusion für die Encoded-Datasets? Im Moment werden die nicht besprochen
% Methode zum CD-Diagramm genau erklären (paired-t-test etc.) und die Relevanz fürs paper zeigen
% Advantage-Matrix von Ji et. al. neu aufarbeiten. Wo sind die Unterschiede? Ist das CD-Diagramm besser/schlechter?
% (Un-)Encoded durch Normal uns Semi-Supervised ersetzen, um die Verbindung klar zu machen
% Eventuell Batchsize 1 von den anderen Ergebnissen separieren. Der Vergleich macht mehr Sinn, als einfach die Batchsizes zu mitteln
% Tabelle 1 für Algorithmen statt Domänen. Das zeigt, dass immer entweder BADGE oder Margin (unsere Top 2) fehlt
% Unsere Rankings mit Literatur vergleichen. Wie anders sind wir? -> Evtl. besten Alg. pro Benchmark sammeln und vergleichen
% Causalitäten checken: Viele Dinge werden einfach genannt, ohne das der Grund richtig eingeführt wird
% Fig. 1: Schlechte Performance am Anfang verwirrt nur. Entweder BADGE benutzen, anders plotten, oder im Text in Kontext setzen
% Oracle Forecast im Appendix erklären

%%%%%%%%%%%%%%%%%%
% DONE
%%%%%%%%%%%%%%%%%
% !! Ergebnis-Tabelle mit nur 3 runs neu berechnen und Unterschiede zeigen
% !! Least-Confident Baseline
% Conclusion: Es gibt Unterschiede bei der Top 1 zwischen den Domänen, aber die Top 3 ist konstant
% !! Sec 3 Related Work sollte Ji et al. im Detail einführen, weil wir uns an deren Guidelines halten. Alles andere sollte dort gekürzt werden
% !! "Overfitting on test" für Oracle: Sanity-Check durchführen und eventuell komplett streichen
% !! Introduction neu schreiben: Wir brauchen eine klare Benennung der Ziele und bessere Abgrenzung zu anderen papern 
% Infos und Ergebnisse für die syntetischen Datensätze in ein Kapitel bündeln und direkt auf die Contribution beziehen
% !! Sec 4 Methodology ist sehr chaotisch. Wir müssen neu sortieren und unwichtige Dinge streichen
% Contributions neu sortieren und ausführlicher beschreiben
% Oracle Performance mit einer Zahl benennen: z.B. Da wird ~20% lift zu allen Baselines haben, brauchen wir kein aufwendigeres Oracle
% "reproducible" als Aufhänger für 50 repeats ist nicht intuitiv
% Eventuell Honeypot und Diverging Sine direkt in den Contributions nennen
% Tabelle mit AUCs in den Appendix

@@

@@

To sum up, \emph{ALBench} is an experimental framework which includes an
efficient oracle approximation, multiple application domains, enough repetitions to
draw valid conclusions and two synthetic tasks to highlight shortcomings of AL methods. 
By being the first benchmark to proving these things in one code-base, 
we believe that \emph{ALBench} is a major step forward of assessing the overall state of Active Learning research, independent of
specific application domains. \emph{ALBench} is publicly available under \emph{anonymous}. \\ [1mm]
Our contributions include the following:
%
\begin{enumerate}
\item\label{cont:repetitions} We show that the small of repetitions that previous works have employed is not
sufficient for meaningful conclusions. Sometimes even making it impossible to assess if a performance is above or below random.
\item\label{cont:oracle} We propose an efficient and performant oracle which is constructed iteratively in a greedy fashion,
overcoming major computational hurdles.
% one of the major obstacles for sufficient repetitions of AL experiments. 
% [Thorben] Das können wir so nicht sagen, glaube ich. Niemand verlangt von uns, dass wir das Oracle auch 50 mal wiederholen
\item\label{cont:domains} We propose \emph{ALBench}, the first general benchmark providing tasks
in the domains of images, text and tabular learning. It further contains
synthetic and pre-encoded data to allow for a sophisticated evaluation of AL
methods. Our experiments show, that there is no clear state-of-the art method
for active learning across different domains.
\item\label{cont:synthdata} We propose \emph{Honeypot} and \emph{Diverging Sin}, two synthetic datasets
designed to hinder active learning by naive decision-boundary- or
clustering-based approaches respectively. Thus, they provide an important tool to identify 
shortcomings of existing AL methods.
\end{enumerate}
%
\begin{table}
	\centering
	\caption{Comparison of our benchmark with the existing literature. Oracle curves
	serve as an approximation of the best possible AL algorithm. Including the
	encoded versions of our datasets we reach 14 datasets, without we have 9. "Semi"
	indicates whether the paper is employing any form of self- or semi-supervised
	learning. A "-" for repetitions means that we could not determine how often each
	experiment is repeated in the respective framework. \emph{ALBench} is the only
	benchmark which reports over enough runs for meaningful results and considers
	all 5 domains.}
	\label{tab:benchmark_comparison}
	\resizebox{\columnwidth}{!}{
\begin{tabular}{lccc|cllll|cl}
Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
\multicolumn{1}{l|}{Beck et al. \cite{beck2021effective}} & batch & 4 & 7 & \checkmark & - & - & - & - & - & - \\
\multicolumn{1}{l|}{Hu et al. \cite{hu2021towards}} & batch & 5 & 13 & \checkmark & \checkmark & - & - & - & - & 3 \\
\multicolumn{1}{l|}{Zhou et al. \cite{zhou2021towards}} & batch & 3 & 2 & \checkmark & \checkmark & - & - & - & \checkmark & 5 \\
\multicolumn{1}{l|}{Zhan et al. \cite{zhan2021comparative}} & sngl+batch & 35 & 18 & - & - & \checkmark & \checkmark & - & \checkmark & 10-100 \\
\multicolumn{1}{l|}{Munjal et al. \cite{munjal2022towards}} & batch & 2 & 8 & \checkmark & - & - & - & - & - & 3 \\
\multicolumn{1}{l|}{Li et al. \cite{li2022empirical}} & batch & 5 & 13 & \checkmark & - & - & - & \checkmark & - & - \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
While multiple benchmark suites have been proposed for Active Learning, none of them provide experiments for more than two domains.
The authors of \cite{beck2021effective}, \cite{munjal2022towards}, \cite{li2022empirical}, \cite{ji2023randomness} and \cite{luth2024navigating} even focus exclusively on the image domain.
Especially the tabular domain is underrepresented in preceding benchmarks, as only \cite{zhan2022comparative} provides experiments for it.
Experiments on the interplay between AL and semi-supervised learning are similarly under-researched, as only two works exist \cite{li2022empirical, luth2024navigating}, both of them only using images.
An oracle algorithm has been proposed by two works \cite{zhou2021towards, zhan2022comparative}. 
Both of these algorithms rely on search and are computationally very expensive, while our proposed method efficiently can be constructed sequentially.
%While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training, and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning.
%The authors of \cite{hu2021towards} study models that are trained on actively learned datasets in the image and text domain.
%They test for several different properties of the models including robustness, response to compression techniques and final performance.
%\cite{zhou2021towards} proposed an oracle algorithm for AL that uses Simulated Annealing search to approximate a solution for the optimal subset of labeled data.
%Additionally, they study the generalization behavior of subsets of labeled data in the text an image domain.
The two closest related works to this benchmark are \cite{ji2023randomness} and \cite{luth2024navigating}, who also place a much higher emphasis on the problem of evaluating AL algorithms under high variance than their predecessors (indicated in Tab. \ref{tab:benchmark_comparison} by a dashed line).
The authors of \cite{ji2023randomness} posed a total of 12 "recommendations" for reliable evaluation of AL algorithms.
We largely adapt the proposed recommendations and extend their work to multiple domains and batch sizes.
For a complete list of the recommendations and our implementation of them, please refer to App. \ref{app:recommendations}.
This work also pays attention to the so-called "pitfalls" of AL evaluation proposed in \cite{luth2024navigating}.
For a complete list of the pitfalls and our considerations regarding them, please refer to App. \ref{app:pitfalls}.
To the best of our knowledge, we are the first to extend reliable SOTA (based on \cite{ji2023randomness, luth2024navigating}) experimentation to a total of 5 data domains and a high number of restarts per experiment.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Few Repetitions are not Sufficient for Meaningful Results}\label{sec:restarts}
To evaluate how many restarts are necessary to obtain conclusive results in an AL experiment, we computed 100 runs of our top-performing algorithm on one dataset.
Our best algorithm is margin sampling and we chose the Splice dataset for its average size and complexity. \\
This allows us firstly, to obtain a very strong estimation of the "true" average performance of margin sampling on this dataset and secondly, to draw subsets from this pool of 100 runs.
Setting the size of our draws to $\alpha$ and sampling uniformly, we can approximate a cross-validation process with $\alpha$ restarts.
Each of these draws can be interpreted as a \textbf{reported result in AL literature} where the authors employed $\alpha$ restarts.
%We provide evidence in Fig. \ref{fig:restarts} that previous works might not have evaluated their experiments with a sufficient number of restarts.
Figure \ref{fig:restarts} shows the "true" mean performance of margin sampling (green) in relation to random sampling (black) and the oracle performance (red).
We display 5 random draws of size $\alpha$ in blue.
We can observe that even for a relatively high number of restarts the variance between the samples is extremely high, resulting in some performance curves being worse that random and some being significantly better.
When setting $\alpha = 50$ we observe all samples to converge close to the true mean performance. 
\begin{figure}
        \hspace{-20mm}
	\includegraphics[width=1.24\linewidth]{img/ablation_restarts}
	\caption{Random draws from a pool of 100 runs for margin sampling on the Splice dataset with different numbers of repetitions ($\alpha=\{5,10,50\}$). Green curves are the mean performance of all 100 runs, while the samples are blue. Even with 5 or 10 repetitions, we can observe that single draws for margin sampling display below-random performance (black), while the true mean should be above random.}
	\label{fig:restarts}
\end{figure}
In addition to this motivating example, we carried out our main evaluation (Tab. \ref{tab:results}) multiple times by uniformly sampling 3 random from our available runs and comparing the results.
We found significant differences in the performance of acquisition functions on individual datasets, as well as permutations in the final ranking.
This partly explains the ongoing difficulties in reproducing results for AL experiments and benchmarks.
The details can be found in App. \ref{app:rank_difference}.
For this benchmark we employ 50 restarts of every experiment.
\begin{wraptable}{R}{0.65\textwidth}
	\vspace{-0.7cm}
	\caption{Employed model, chosen budget and available batch sizes for each dataset}
	\vspace{0.1cm}
	\label{tab:batch_sizes}
	{\scriptsize
		\begin{tabular}{l|l|c|c|c|c|c|c|c|c}
			& Model  & B & 1 & 5 & 20 & 50 & 100 & 500 & 1K \\
			\hline
			Enc. DNA    & Linear & 40 & o & o &&&&& \\
			Enc. Splice & Linear & 100 & o & o & o & o &&&\\
			TopV2       & BiLSTM & 200 & o & o & o & o &&& \\
			Splice      & MLP    & 400 & o & o & o & o & o && \\
			DNA         & MLP    & 300 & o & o & o & o & o && \\
			USPS        & MLP    & 400 & o & o & o & o & o && \\
			Enc. Cifar10& Linear & 450 & o & o & o & o & o && \\
			Enc. FMnist & Linear & 500 & o & o & o & o & o && \\
			Enc. USPS   & Linear & 600 & o & o & o & o & o && \\
			News        & BiLSTM & 3K &&& o & o & o & o &\\
			FMnist      & ResNet18& 10K &&&&&& o & o\\
			Cifar10     & ResNet18& 10K &&&&&& o & o \\
		\end{tabular}
	}
	\vspace{-0.55cm}
\end{wraptable}
